Comparing Apples and Oranges in Youth Depression Treatments?

### 2024-10-04 multilevel_metaregression/manuscript_render_test.qmd
saving test files

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
subtitle: "A Quantitative Critique of the Evidence Base and Guidelines"
author:
  - name: Argyris Stringaris
    email: a.stringaris@ucl.ac.uk
    affiliations: 
        - id: ucl_1
          name: University College London
          department: Divisions of Psychiatry and Psychology and Language Science
          address: 1-19 Torrington Place
          city: London, UK
          state: 
          postal-code: WC1E 7HB
          country: United Kingdom
    attributes:
        corresponding: true
    note: 
  - name: Charlotte Burman
    email: charlotte.burman@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
  - name: Raphaelle Delpech
    email: raphaelle.delpech.18@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Rudolf Uher
    email: uher@Dal.Ca
    affiliations:
        - id: dal
          name: Dalhousie University
          department: Department of Psychiatry
          address: 
          city: Nova Scotia
          state: 
          postal-code: 
          country: Canada
  - name: Dayna Bhudia
    email: dayna.bhudia@ucl.ac.uk
    affiliations:
        - ref: ucl_1
  - name: Despoina Miliou
    email: dgmiliou@outlook.com.gr
    affiliations:
        - ref: ucl_1
  - name: Ioannis-Marios Rokas
    email: rokasgiannis93@gmail.com
    affiliations:
        - ref: ucl_1
  - name: Marinos Kyriakopoulos
  - name: Lucy Foulkes
    email: lucy.foulkes@psych.ox.ac.uk
    affiliations:
        - id: OU
          name: University of Oxford
          department: Experimental Psychology
          address: Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Road
          city: Oxford
          postal-code: OX2 6GG
          country: United Kingdom
    note:
  - name: Carmen Moreno
    email: cmoreno@hggm.es
    affiliations:
        - id: HGM
          name: Hospital Gregorio Marañón
          department: Department of Psychiatry
          address: 46 C. del Dr. Esquerdo
          city: Madrid
          state:
          postal-code: 28007
          country: Spain
  - name: Samuele Cortese
    email: samuele.cortese@soton.ac.uk
    affiliations:
        - id: soton
          name: University of Southampton
          department: Centre for Innovation in Mental Health (CIMH), School of Psychology
          address: Highfield Campus, Building 44
          city: Southampton
          state:
          postal-code: SO17 1BJ
          country: United Kingdom
  - name: Glyn Lewis
    email: glyn.lewis@ucl.ac.uk
    affiliations:
        - ref: ucl_1        
  - name: Georgina Krebs
    email: g.krebs@ucl.ac.uk
    affiliations:
        - ref: ucl_1
    note:
keywords: 
  - depression
  - adolescents
  - antidepressants
  - psychotherapy
date: last-modified
format:
  docx:
    reference-doc: "bibliography/word-style-reference-doc.docx"
    fig-width: 9
    fig-height: 7
    fig_caption: yes
    tbl_caption: yes
    number-sections: true
execute: 
  echo: true
  auto: true
  warning: FALSE
bibliography: "bibliography/references.bib"
csl: "bibliography/vancouver_style.csl"
font:
  family: Calibri
fig-cap-location: top
---

@@

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)
library(here)

@@

# For referencing
bib_data <- ReadBib(here("multilevel_metaregression", "bibliography", "references.bib"))

@@

**Objectives:** Should a young person receive psychotherapy or medication for their depression, and on what evidence do we base this decision? In this paper, we test whether the basic conditions required to draw valid inferences to answer this question are currently met. **Methods:** We included 88 RCTs of psychotherapy and medication for child and adolescent depression (mean age 4-18 years). Using meta-analyses, we compared a) participant characteristics and b) trial characteristics in medication and psychotherapy trials. Lastly, we examined whether psychotherapy controls are well-matched to active conditions. **Results:** Participants in medication RCTs had higher depression severity and were more frequently male compared to psychotherapy RCTs. There was a dramatic difference in the within-subject improvement due to placebo (SMD=-1.9 (95% CI: -2.10 to -1.70)) vs psychotherapy controls (SMD=-0.5 (95% CI: -0.75 to -0.25)). Within psychotherapy RCTs, control conditions were less intensive on average than active conditions. **Conclusions:** Medication and psychotherapy RCTs differ on fundamental participant and methodological characteristics, thereby violating key conditions for valid comparison between them. Psychotherapy controls often involve little therapist contact and are easy-to-beat comparators. These findings cast doubt on the confidence with which psychotherapy is recommended for youth depression, and highlight the pressing need to improve the evidence base.

@@

```{r define-functions, warning=FALSE, message = FALSE, echo=FALSE}
#############Some custom made functions#############

@@

  df_coefs <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_se <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_z_value <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_lower_ci <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_upper_ci <- matrix(NA, nrow = n_models, ncol = n_conditions)
  tau_sq_1 <- rep(NA, n_models)
  tau_sq_2 <- rep(NA, n_models)
  qe <- rep(NA, n_models)
  k <- rep(NA, n_models)

  for (i in seq_len(n_models)) {
    model <- list_model_1_meta_reg[[i]]
    tau_sq_1[i] <- model$sigma2[1]
    tau_sq_2[i] <- model$sigma2[2]
    qe[i] <- model$QE
    k[i] <- model$k
    coefs <- model$beta
    ses <- model$se
    zvals <- model$zval
    lower_ci <- model$ci.lb
    upper_ci <- model$ci.ub

@@

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs, na.rm = TRUE),
    se = colMeans(df_se, na.rm = TRUE),
    z_value = colMeans(df_z_value, na.rm = TRUE),
    lower_ci = colMeans(df_lower_ci, na.rm = TRUE),
    upper_ci = colMeans(df_upper_ci, na.rm = TRUE),
    tau_sq_1 = mean(tau_sq_1, na.rm.= TRUE),
    tau_sq_2 = mean(tau_sq_2, na.rm.= TRUE),
    qe = mean(qe, na.rm = TRUE),
    k = mean(k, na.rm = TRUE)
  )

@@

###########8. Function for extracting multilevel I2 estimates

extract_i2_values <- function(list_of_models) {
  # Initialize vectors to store I2 values
  i2_level2 <- numeric(length(list_of_models))
  i2_level3 <- numeric(length(list_of_models))

  # Loop through each fitted model from the list
  for (i in seq_along(list_of_models)) {
    model <- list_of_models[[i]]

    # Calculate variance distribution using the existing function
    i2_results <- var.comp(model)

    # Extract I2 values for level 2 and level 3 from the results
    i2_level2[i] <- as.numeric(i2_results$results$I2[2]) # I2 for level 2
    i2_level3[i] <- as.numeric(i2_results$results$I2[3]) # I2 for level 3
  }

  # Combine I2 values into a data frame
  i2_data <- data.frame(
    i2_level2 = i2_level2,
    i2_level3 = i2_level3
  )

  return(i2_data)
}

@@

```{r import-dataset, warning=FALSE, message = FALSE, echo=FALSE}

@@

df_appl_v_orange <-readxl::read_excel(here("df_appl_v_orange.xlsx"))

@@

df_scale_ranges <- read.csv(here("scale_ranges.csv"))
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

```{r metaregression-all, warning=FALSE, message = FALSE, echo=FALSE}

@@

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the CBT vs Fluox and Escitalopram sample
# C) the no WL sample
# D) clinical levels of depression
# E) the CDRS only sample
# E) HAMD sensitivity analysis
# G) Studies with post SD
# H) Studies with variance <.02

@@

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

# Extract I2 values for each dataset
i2_results_all <- extract_i2_values(list_model_1_meta_reg)

# Calculate the mean and standard deviation of I2 values for each level
i2_summary_all <- data.frame(
  mean_i2_level2 = mean(i2_results_all$i2_level2, na.rm = TRUE),
  sd_i2_level2 = sd(i2_results_all$i2_level2, na.rm = TRUE),
  mean_i2_level3 = mean(i2_results_all$i2_level3, na.rm = TRUE),
  sd_i2_level3 = sd(i2_results_all$i2_level3, na.rm = TRUE)
)
```

```{r metaregression-fluox, warning=FALSE, message = FALSE, echo=FALSE}
# B. Estimate Means and CIs for CBT Fluox Esc --------------------------------
#  grab ids for those studies
cbt_fluox_esc_study_ids <-
  df_appl_v_orange[df_appl_v_orange$active_type == "cbt" | df_appl_v_orange$active_type == "Fluoxetine" | df_appl_v_orange$active_type == "Escitalopram",]$new_study_id

# Use those ids to subset all dataframes in the list
# By searching for IDs we have mistakenly included multiple active arms. E.g. for Atkinson which has dulox and fluox.
# So we need to further subset the simulated dfs. For the active arms, I am filtering out conditions that do not meet our criteria.
# This leaves all the control conditions from trials where the active is cbt, fluox or esc.
# We may have multiple controls per study id for this reason, though this isn't a problem
# conceptually.

list_cbt_fluox_esc_study <- lapply(list_df_simulated, function(df) {
  df %>%
    filter(new_study_id %in% cbt_fluox_esc_study_ids) %>%
    filter(!(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
})

# count the number of studies
cbt_fluox_esc_study <- df_long_for_metan %>%
  filter(new_study_id %in% cbt_fluox_esc_study_ids,
         !(arm_effect_size == "cohens_d_active" & !(type %in% c("Fluoxetine", "Escitalopram", "cbt"))))
study_count_cbt_fluox_esc_study <- count_studies(cbt_fluox_esc_study)
# CB checked, looks good now.

# apply the metareg function to the list of cbt, fluox, esc
model_1 <- as.formula(~ four_level_var)
list_model_1_cbt_fluox_esc_study <- run_multi_metaregression(list_cbt_fluox_esc_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_cbt_fluox_esc_study <-  aggregate_model_results(list_model_1_cbt_fluox_esc_study, condition)
aggregate_results_cbt_fluox_esc_study <- cbind(aggregate_results_cbt_fluox_esc_study, condition)

# extract SMDs from the list for cbt fluox escit
means_cbt_fluox_esc_study <-  lapply(list_model_1_cbt_fluox_esc_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_cbt_fluox_esc_study <- calculate_mean_coefs_ses(means_cbt_fluox_esc_study, condition)
 coef_and_se_means_cbt_fluox_esc_study <- data.frame(cbind(condition, coef_and_se_means_cbt_fluox_esc_study))

# add the ns to the dataframe
coef_and_se_means_cbt_fluox_esc_study <- left_join(coef_and_se_means_cbt_fluox_esc_study, study_count_cbt_fluox_esc_study[study_count_cbt_fluox_esc_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")
```

```{r metaregression-nowl, warning=FALSE, message = FALSE, echo=FALSE}

# C. Estimate Means and CIs without WL ---------------------------------------

@@

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-clin, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-cdrs, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
```

```{r metaregression-hamd, warning=FALSE, message = FALSE, echo=FALSE}

@@

# add the ns to the dataframe
coef_and_se_means_hamd_study <- left_join(coef_and_se_means_hamd_study, study_count_hamd_study[study_count_hamd_study$is_missing == FALSE, ] %>% 
                                        ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")

@@

```{r metaregression-postsd, warning=FALSE, message = FALSE, echo=FALSE}

# G. Post SD Sensitivity Analysis ---------------------------------------

##--------- Data preparation

#For this we need to filter the master df by missing post_sd_active and post_sd_control in df_first_row

#create subset of first_row

df_first_row <- read.csv(here("df_first_row.csv"))

#create new_study_ids to match master dataset
df_first_row  <- df_first_row  %>%
  mutate(new_study_id = case_when(psy_or_med == 0 ~ paste(study,year, sep = ", "),
                                  .default = study ))

df_first_row_reduced <- df_first_row %>%
  filter(!(new_study_id == "March, 2004" & is.na(control_type))) %>%
  select(new_study_id, post_sd_active, post_sd_control)

#as first_row has multiple rows per study (showing different arm pairs), I will keep only one row per study

df_first_row_reduced <- df_first_row_reduced %>%
  group_by(new_study_id) %>%
  slice_head(n = 1) %>%
  ungroup()

#create a column in master df that shows 1 if any missing post sds in first_row

df_long_for_metan_with_missing <- df_long_for_metan %>%
  left_join(df_first_row_reduced, by = "new_study_id") %>%
  mutate(missing_post_sd = ifelse(is.na(post_sd_active) | is.na(post_sd_control), 1, 0)) %>%
  select(-post_sd_active, -post_sd_control) %>%
  arrange(arm_effect_size, new_study_id)

#all studies missing sds in active are also missing them in control, so we can go ahead and filter out all arms of these studies

df_long_for_metan_without_missing <- df_long_for_metan_with_missing %>%
  filter(missing_post_sd == 0)

n_distinct(df_long_for_metan_without_missing$new_study_id) #61 studies
n_distinct(df_long_for_metan$new_study_id) #88 studies total
nrow(df_long_for_metan_without_missing) #140 arms
nrow(df_long_for_metan) #203 arms

##------------ Estimates

#take study ids where study does not have missing post SD
sd_study_ids <-
  df_long_for_metan_with_missing[df_long_for_metan_with_missing$missing_post_sd ==0,]$new_study_id

# use those ids to subset all dataframes in the list
list_sd_study <- list()

  list_sd_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                                 %in% sd_study_ids,]
}

# count the number of studies
study_count_sd_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% sd_study_ids,])

# apply the metareg function to the list of clinical studies
model_1 <- as.formula(~ four_level_var)
list_model_1_sd_study <- run_multi_metaregression(list_sd_study , model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_sd_study <-  aggregate_model_results(list_model_1_sd_study, condition)

# extract SMDs from the list for clinical studies
means_sd_study <-  lapply(list_model_1_sd_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_sd_study <- calculate_mean_coefs_ses(means_sd_study, condition)
coef_and_se_means_sd_study <- data.frame(cbind(condition, coef_and_se_means_sd_study))

# add the ns to the dataframe
coef_and_se_means_sd_study <- left_join(coef_and_se_means_sd_study, study_count_sd_study[study_count_sd_study$is_missing == FALSE, ] %>%
                                            ungroup() %>%
                                            select(condition, n),
                                          by = "condition")

```{r metaregression-var, warning=FALSE, message = FALSE, echo=FALSE}

#H. Studies with variances <.02-----------------------------

##-----------Data preparation

#create column with variance in master df
df_long_for_metan$baseline_variances <- (df_long_for_metan$perc_baseline_sd)^2

#take study ids where variance =2 or less
var2_study_ids <-
  df_long_for_metan[df_long_for_metan$baseline_variances<=0.02,]$new_study_id
length(var2_study_ids)

##------------Estimates

# use those ids to subset all dataframes in the list
list_var2_study <- list()

  list_var2_study[[i]] <- list_df_simulated[[i]][list_df_simulated[[i]]$new_study_id
                                               %in% var2_study_ids,]
}

# count the number of studies
study_count_var2_study <- count_studies(df_long_for_metan[df_long_for_metan$new_study_id %in% var2_study_ids,])

# apply the metareg function to the list
model_1 <- as.formula(~ four_level_var)
list_model_1_var2_study <- run_multi_metaregression(list_var2_study, model_1)

# extract coefficients and model characteristics.
condition <- levels(list_df_simulated[[1]]$four_level_var)
aggregate_results_var2_study <-  aggregate_model_results(list_model_1_var2_study, condition)

# extract SMDs from the list
means_var2_study <-  lapply(list_model_1_var2_study, extract_coefficients_func)

# calculate mean SMDs and ses
condition <- levels(list_df_simulated[[1]]$four_level_var) # for the conditino argument
coef_and_se_means_var2_study <- calculate_mean_coefs_ses(means_var2_study, condition)
coef_and_se_means_var2_study <- data.frame(cbind(condition, coef_and_se_means_var2_study))

# add the ns to the dataframe
coef_and_se_means_var2_study <- left_join(coef_and_se_means_var2_study, study_count_var2_study[study_count_var2_study$is_missing == FALSE, ] %>%
                                          ungroup() %>%
                                          select(condition, n),
                                        by = "condition")
```

@@

```{r baseline-severity, echo=FALSE, message=FALSE, warning=FALSE}

mean_psych_overall <- 0.37
mean_med_overall <- 0.42
se_psych_overall <- 0.02
se_med_overall <- 0.01
n_psych_overall <- 49
n_med_overall <- 31

# Calculate standard deviation
sd_psych_overall <- se_psych_overall * sqrt(n_psych_overall)
sd_med_overall <- se_med_overall * sqrt(n_med_overall)

#formula with pooled sd
pooled_sd_overall <- sqrt(((sd_psych_overall^2 + sd_med_overall^2)/2))

# Calculate Cohen's d
cohens_d_overall <- (mean_psych_overall - mean_med_overall) / pooled_sd_overall

## without subclinical

mean_psych_sub <- 0.39
mean_med_sub <- 0.42
se_psych_sub <- 0.02
se_med_sub <- 0.01
n_psych_sub <- 41
n_med_sub <- 31

# Calculate standard deviation
sd_psych_sub <- se_psych_sub * sqrt(n_psych_sub)
sd_med_sub <- se_med_sub * sqrt(n_med_sub)

#formula with pooled sd
pooled_sd_sub <- sqrt(((sd_psych_sub^2 + sd_med_sub^2)/2))

# Calculate Cohen's d
cohens_d_sub <- (mean_psych_sub - mean_med_sub) / pooled_sd_sub
cohens_d_sub
cohens_d_overall
```

@@

To ensure that this was not an artefact of variable transformation, we also compared means at baseline in the two instruments, CDRS and HAM-D, on which there was a sufficient number of studies to metanalyse. As can be seen in @tbl-hamd-baseline and @tbl-cdrs-baseline (see Supplemental Materials), the number of studies is much smaller, but the pattern of differences is the same for the HAM-D and the CDRS, though it does not reach statistical significance for the latter.

@@

For this analysis, we excluded the two psychotherapy trials which included entirely female samples (Moeini, 2019; Shomaker, 2016; see @tbl-all-studies-percfemale-results in Supplement for results with full sample). As can be seen in @tbl-baseline_results, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

@@

As can be seen in tbl-baseline_results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = ``` r``round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2) ```) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

plot_smd_diff_control <- list_df_simulated[[1]] %>%
  filter(arm_effect_size == "cohens_d_control") %>%
  ggplot(aes(x = cohens_d, y = 0, color = as.factor(psy_or_med), size = baseline_n)) +
  geom_jitter(height = 0.2, alpha = 0.7) +
  geom_point(alpha = 0.7) +
  labs(
    x = "SMD",
    y = "",
    color = "Type of Study",
    size = "Baseline n"
  ) +
  scale_color_discrete(name = "Type of Study", labels = c("Medication", "Psychological Therapy")) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )

print(plot_smd_diff_control)

png("plot_smd_diff_control.png", width=6.5, height=4, units="in", res=300)
 plot_smd_diff_control
 dev.off()

@@

 png("plot_means_multi_all.png", width=6.5, height=6, units="in", res=300)
 plot_means_multi_all
 dev.off()

```{r, warning=FALSE, message = FALSE, echo=FALSE, ft.align="left"}
# #| label: tbl-coefs-overall
# #| tbl-cap: "Summary statistics of the estimated SMDs"
# #| tbl-cap-location: top
# 
# # Prepare to tabulate
# 
# table_mean_coefs_from_sim <- df_mean_coefs_from_sim %>%
#   mutate(across(where(is.numeric), ~round(., 2))) %>%
#   relocate("n", .after = "condition") 
# 
# new_column_names <- c("Condition", "N", "Coefficient", "SE", "Lower CI", "Upper CI")
# 
# # Assign new column names to the data frame
# colnames(table_mean_coefs_from_sim) <- new_column_names
# 
# flextable(table_mean_coefs_from_sim) %>%
#   set_table_properties(width = 1) %>%
#   autofit() %>%
#   font(fontname = "Calibri", part = "all") %>% 
#   # add_header_row(values = "Summary statistics of the estimated SMDs", colwidths = 6) %>%
#   # add_header_row(values = "Table 3", colwidths = 6) %>% 
#   bold(part = "header") %>%   # Bold the relevant header rows
#   bg(bg = "#F2F2F2", part = "header") %>%   # Add light grey background to header row %>% 
#   # italic(i = 2, italic = TRUE, part = "header") %>% 
#   border_remove() %>% # this is to remove borders around the caption that look strange
#   hline(part = "body", border = small_border) %>%  # Add light grey horizontal lines all  rows
#   vline(j = 1, part = "body", border = small_border) %>% 
#   border_outer(border = small_border, part = "body" ) %>%  # Add outer borders
#   hline(part = "header", border = big_border) %>% 
#   # padding(i = 2, padding.bottom = 15, part = "header") %>% 
#   align(j = 2:ncol(df_mean_coefs_from_sim), align = "center", part = "all") %>% 
#   align(j = 1, align = "left", part = "all") %>% 
#   paginate(init = TRUE, hdr_ftr = TRUE) %>% 
#   line_spacing(space = .7, part = "all")

@@

>>>>

---
title: "Comparing Apples and Oranges in Youth Depression Treatments?"
author: "Argyris Stringaris"
date: "05-09-2024"
format:
  docx:
    reference-doc: "bibliography/word-style-reference-doc.docx"
    fig-width: 9
    fig-height: 7
    fig_caption: true
    tbl_caption: true
    number-sections: true
execute: 
  echo: true
  warning: FALSE
bibliography: "bibliography/references.bib"
csl: "bibliography/vancouver_style.csl"
font:
  family: Calibri
fig-cap-location: top
---

@@

###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
library(RefManageR)
library(ftExtra)
library(dplyr)
library(broom)
library(effsize)
library(here)
library(officer)

@@

# For referencing
bib_data <- ReadBib(here::here("multilevel_metaregression", "bibliography", "references.bib"))

#some set-up for flextables
big_border = fp_border(color="black", width = 1)
small_border = fp_border(color="lightgray", width = 1)
```

@@

**Objectives:** Should a young person receive psychotherapy or medication for their depression, and on what evidence do we base this decision? In this paper, we test whether the basic conditions required to draw valid inferences to answer this question are currently met.

**Methods:** We included 88 RCTs of psychotherapy and medication for child and adolescent depression (mean age 4-18 years). Using meta-analyses, we compared a) participant characteristics and b) trial characteristics in medication and psychotherapy trials. Lastly, we examined whether psychotherapy controls are well-matched to active conditions.

**Results:** Participants in medication RCTs had higher depression severity and were more frequently male compared to psychotherapy RCTs. There was a dramatic difference in the within-subject improvement due to placebo (SMD=-1.9 (95% CI: -2.10 to -1.70)) vs psychotherapy controls (SMD=-0.5 (95% CI: -0.75 to -0.25)). Within psychotherapy RCTs, control conditions were less intensive on average than active conditions.

**Conclusions:** Medication and psychotherapy RCTs differ on fundamental participant and methodological characteristics, thereby violating key conditions for valid comparison between them. Psychotherapy controls often involve little therapist contact and are easy-to-beat comparators. These findings cast doubt on the confidence with which psychotherapy is recommended for youth depression, and highlight the pressing need to improve the evidence base.

@@

```{r define-functions, warning=FALSE, message = FALSE, echo=FALSE, cache=FALSE}
#############Some custom made functions#############

@@

  df_coefs <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_se <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_z_value <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_lower_ci <- matrix(NA, nrow = n_models, ncol = n_conditions)
  df_upper_ci <- matrix(NA, nrow = n_models, ncol = n_conditions)
  tau_sq <- rep(NA, n_models)
  qe <- rep(NA, n_models)
  k <- rep(NA, n_models)

  for (i in seq_len(n_models)) {
    model <- list_model_1_meta_reg[[i]]
    tau_sq[i] <- model$tau2
    qe[i] <- model$QE
    k[i] <- model$k
    coefs <- model$beta
    ses <- model$se
    zvals <- model$zval
    lower_ci <- model$ci.lb
    upper_ci <- model$ci.ub

@@

  df_coefficients_model <- data.frame(
    coefficients = colMeans(df_coefs, na.rm = TRUE),
    se = colMeans(df_se, na.rm = TRUE),
    z_value = colMeans(df_z_value, na.rm = TRUE),
    lower_ci = colMeans(df_lower_ci, na.rm = TRUE),
    upper_ci = colMeans(df_upper_ci, na.rm = TRUE),
    tau_sq = mean(tau_sq, na.rm = TRUE),
    qe = mean(qe, na.rm = TRUE),
    k = mean(k, na.rm = TRUE)
  )

@@

@@

```{r import-dataset, warning=FALSE, message = FALSE, echo=FALSE, cache=FALSE}

@@

df_appl_v_orange <-readxl::read_excel(here::here("df_appl_v_orange.xlsx"))

@@

df_scale_ranges <- read.csv(here::here("scale_ranges.csv"))
df_appl_v_orange <- full_join(df_appl_v_orange, df_scale_ranges, by = "instrument_name")

@@

```{r run-metaregressions, cache=FALSE, warning=FALSE, message = FALSE, echo=FALSE}

@@

######## NOW RUN METAREGRESSIONS AND GET Coefficients for 
# A) the overall sample
# B) the no WL sample
# C) clinical levels of depression
# D) the CDRS only sample
# E) HAMD sensitivity analysis
# F) Studies with post SD
# G) Studies with variance <.02

@@

# add the ns
df_mean_coefs_from_sim <- left_join(df_mean_coefs_from_sim, df_count_studies[df_count_studies$is_missing == FALSE, ] %>% 
                                      ungroup() %>% 
                                      select(condition, n),  
                                    by = "condition")
# C. Estimate Means and CIs without WL---------------------------------------

@@

coef_and_se_means_no_wl <- left_join(coef_and_se_means_no_wl, study_count_no_wl_study[study_count_no_wl_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")

@@

# add the ns to the dataframe
coef_and_se_means_clin_study <- left_join(coef_and_se_means_clin_study, study_count_clin_study[study_count_clin_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")

@@

# add the ns to the dataframe
coef_and_se_means_cdrs_study <- left_join(coef_and_se_means_cdrs_study, study_count_cdrs_study[study_count_cdrs_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")

@@

# add the ns to the dataframe
coef_and_se_means_hamd_study <- left_join(coef_and_se_means_hamd_study, study_count_hamd_study[study_count_hamd_study$is_missing == FALSE, ] %>%
                                        ungroup() %>%
                                      select(condition, n),
                                    by = "condition")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#| label: tbl-all-trials
#| tbl-cap: "Summary of included RCTs"
#| tbl-cap-location: top

@@

@@

@@

To ensure that this was not an artefact of variable transformation, we also compared means at baseline in the two instruments, CDRS and HAM-D, on which there was a sufficient number of studies to metanalyse. As can be seen in @tbl-hamd-baseline and @tbl-cdrs-baseline (see Supplement), the number of studies is much smaller, but the pattern of differences is the same for the HAM-D and the CDRS, though it does not reach statistical significance for the latter.

@@

For this analysis, we excluded the two psychotherapy trials which included entirely female samples (Moeini, 2019; Shomaker, 2016; see @tbl-all-studies-percfemale-results in Supplement for results with full sample). As can be seen in @tbl-baseline-results, psychotherapy trials featured a significantly higher percentage of females when compared to medication trials. On average, samples were `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 1][[2]],2)`) female across psychotherapy trials and `r round(results_met_perc_women_overall$TE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`% (*SE* = `r round(results_met_perc_women_overall$seTE_random[results_met_perc_women_overall$subgroup == 0][[2]],2)`) female across medication trials. Excluding sub-clinical and waitlist control studies yielded similar results.

@@

As can be seen in @tbl-baseline-results, mean age was `r round(results_met_age$TE_random[results_met_age$subgroup == 1][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 1][[2]],2)`) across psychotherapy trials and `r round(results_met_age$TE_random[results_met_age$subgroup == 0][[2]],2)` (*SE* = `r round(results_met_age$seTE_random[results_met_age$subgroup == 0][[2]],2)`) across medication trials, with no significant between group differences. There were no significant differences in mean age between modalities on further sensitivity analyses.

@@

list_df_simulated[[1]] %>%
  filter(arm_effect_size == "cohens_d_control") %>%
  ggplot(aes(x = cohens_d, y = 0, color = as.factor(psy_or_med), size = baseline_n)) +
  geom_jitter(height = 0.2, alpha = 0.7) +
  geom_point(alpha = 0.7) +
  labs(
    x = "SMD",
    y = "",
    color = "Type of Study",
    size = "Baseline n"
  ) +
  scale_color_discrete(name = "Type of Study", labels = c("Medication", "Psychological Therapy")) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank()
  )
```

@@

# png("plot_means_multi_all.png")
# plot_means_multi_all
# dev.off()

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}
#| label: fig-plot-means-no-wl
#| fig-cap: "Meta-analytic estimates of within-group changes: waitlist studies excluded"
#| fig-cap-location: top
#| fig-title: Figure S

 #png("plot_means_multi_no_wl.png")
#plot_means_no_wl
 #dev.off()

```{r, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}
#| label: fig-plot-means-clin
#| fig-cap: "Meta-analytic estimates of within-group changes: subclinical studies excluded"
#| fig-cap-location: top

# png("plot_means_multi_clin.png")
# plot_means_clin
# dev.off()

```{r, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}
#| label: fig-plot-means-cdrs
#| fig-cap: "Meta-analytic estimates of within-group changes: CDRS studies only"
#| fig-cap-location: top

 # png("plot_means_multi_cdrs.png")
 # plot_means_cdrs
 # dev.off()

```{r, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}
#| label: fig-plot-means-hamd
#| fig-cap: "Meta-analytic estimates of within-group changes: HAM-D studies only"
#| fig-cap-location: top

 # png("plot_means_multi_hamd.png")
 # plot_means_multi_hamd
 # dev.off()
